
# Poisson Models

## Overview 

<p style="text-align:justify;">
Poisson regression via ```sgdnet()``` is now supported by **sgdnet**, which has the following objective function in elastic net regularization:
</p>

$$
-\frac1n \sum_{i=1}^n \Big( y_i (\beta_0 + \beta^\mathsf{T} x_i)
- e^{\beta_0 + \beta^\mathsf{T} x_i} \Big)
+ \lambda \left( \frac{1 - \alpha}{2} ||\beta||_2^2
+ \alpha||\beta||_1 \right)
$$
<p style="text-align:justify;">
We use the deviance function for a Poisson distribution $y_i \sim Pois(\mu_i)$ to evaluate model loss:
</p>

$$
2 \sum_{i=1}^n \Big [ y_i log \, (\frac{y_i}{\mu_i}) - (y_i - \mu_i)  \Big ]
$$
where $\mu_i = e^{\beta_0 + \beta^\mathsf{T} x_i}$.

<p style="text-align:justify;">
The Poisson loss function is convex but not Lipschitz-continuous. To guarantee SAGA's convergence, which rely on derivatives' Lipschitz-continuity, we will update the approximated Lipschitz constant $L$ while running the algorithm. Starting with an initial estimate $L^{0}$ as $L_{\text{max}}$, which is the largest sample-wise squared norm of the feature matrix, an invalid $L^{k}$ will be doubled after evaluating $f_i$ with the coefficient $\beta_{k}$ at iteration $k$, if the following inequality is not satisfied [@schmidt2016]:
</p>

$$
f_{i_{k}} \Big ( (\beta_{0k} + \beta_{k}^\mathsf{T} x_i) - \frac{f^{'}_{i_{k}}(\beta_{0k},\beta_{k})}{L^{k}}  \left \| x_i \right \|^2_{2}  \Big ) \leq f_{i_{k}} (\beta_{0k} + \beta_{k}^\mathsf{T} x_i) - \frac{1}{2L^{k}} \left \| f^{'}_{i_{k}}(\beta_{0k}, \beta_{k}) \right \|^2 \left \| x_i \right \|^2_{2}
$$
<p style="text-align:justify;">
where the individual norm is precomputed. The corresponding cross-validation method via ```cv_sgdnet()```, as well as making prediction by ```predict()``` and model performance measure by ```score()``` are also avilable for Poisson in **sgdnet**.
</p>

## Example

<p style="text-align:justify;">
For illustration, we compare with [**glmnet**](https://cran.r-project.org/web/packages/glmnet/index.html) using the fitting results of caddisfly dataset along regularization path. This dataset describes the abundances of one of 17 species in [original dataset](https://pbil.univ-lyon1.fr/ADE-4/ade4-html/trichometeo.html), which is Stactobiella risi, and 12 meteorological features that may influence its presence during 49 trapping night.
</p>

```{r echo = FALSE, message = FALSE}
devtools::load_all("~/sgdnet/family/poisson/sgdnet")
library(glmnet)
```

```{r poisson}
data("caddisfly")
x <- caddisfly$x
y <- caddisfly$y
sfit0 <- sgdnet(x, y, family = "poisson", alpha = 0)
gfit0 <- glmnet(x, y, family = "poisson", alpha = 0)
```

```{r plot, out.width="50%", out.extra='style="; padding:10px; display: inline-block;"', echo=FALSE}
plot(sfit0)
plot(gfit0)
```

<left>
Poisson model with ridge regression on caddisfly dataset
</left>

<br />
```{r poisson2}
sfit1 <- sgdnet(x, y, family = "poisson", alpha = 1)
gfit1 <- glmnet(x, y, family = "poisson", alpha = 1)
```

```{r plot2, out.width="50%", out.extra='style="display: inline-block;"', echo=FALSE}
plot(sfit1)
plot(gfit1)
```

<left>
Poisson model with lasso regression on caddisfly dataset
</left>

# Reference
